---
title: "Written_homework_final_draft_comp"
author: "Jeffrey Ochavillo"
date: "2023-05-01"
output: html_document
---

```{r}
library(tidyverse)
library(e1071)
library(ISLR2)
library(tidyverse)
library(ISLR2)
library(tree)
library(randomForest)
library(dplyr)
library(gbm)
library(dplyr)
library(haven)
```

```{r}
load("/Users/jeffreyochavillo/DATA 5322/written_homework2/Housing.Rdata")
```



```{r}
data1 <- data
```


```{r}
variables <- c('OWNERSHP','INCTOT','AGE','MARST','EDUC','EDUCD','VEHICLES','BEDROOMS')
peoplefactors <- data[,variables]
dim(peoplefactors)
head(peoplefactors)
```

Your task is to predict, based on factors about the people or housing itself, if a dwelling is occupied by owners or renters

is housing/shelter is occupied by owners or renters
So I need OWNERSHP, 

```{r}
#need to filter the data where it's only containing values from those who can actually rent or own a home
#EDUCATION (7 - 11)
#AGE be greater than 18

clean_pf <- peoplefactors %>%
  filter(AGE >= 18) %>%
  filter(EDUC >= 7) %>%
  filter(EDUCD >= 62)

clean_pf

```

```{r}
data$AGE <- zap_labels(data$AGE)
data$AGE <- as.numeric(data$AGE)
data$INCTOT <- zap_labels(data$INCTOT)
data$INCTOT <- as.numeric(data$INCTOT)
data$MARST <- zap_labels(data$MARST)
data$MARST <- as.numeric(data$MARST)
data$EDUC <- zap_labels(data$EDUC)
data$EDUC <- as.numeric(data$EDUC)
data$OWNERSHPD <- zap_labels(data$OWNERSHPD)
data$OWNERSHPD <- as_factor(data$OWNERSHPD)
data$OWNERSHP <- zap_labels(data$OWNERSHP)
data$OWNERSHP <- as_factor(data$OWNERSHP)

#ADD THIS
data$EDUCD <- zap_labels(data$EDUCD)
data$EDUCD <- as.numeric(data$EDUCD)
data$VEHICLES <- zap_labels(data$VEHICLES)
data$VEHICLES <- as.numeric(data$VEHICLES)
```


```{r}
clean_pf$AGE <- zap_labels(clean_pf$AGE)
clean_pf$AGE <- as.numeric(clean_pf$AGE)
clean_pf$INCTOT <- zap_labels(clean_pf$INCTOT)
clean_pf$INCTOT <- as.numeric(clean_pf$INCTOT)
clean_pf$MARST <- zap_labels(clean_pf$MARST)
clean_pf$MARST <- as.numeric(clean_pf$MARST)
clean_pf$EDUC <- zap_labels(clean_pf$EDUC)
clean_pf$EDUC <- as.numeric(clean_pf$EDUC)
clean_pf$OWNERSHPD <- zap_labels(clean_pf$OWNERSHPD)
#clean_pf$OWNERSHPD <- as_factor(clean_pf$OWNERSHPD)
clean_pf$OWNERSHP <- zap_labels(clean_pf$OWNERSHP)
clean_pf$OWNERSHP <- as_factor(clean_pf$OWNERSHP)

#ADD THIS
clean_pf$EDUCD <- zap_labels(clean_pf$EDUCD)
clean_pf$EDUCD <- as.numeric(clean_pf$EDUCD)
clean_pf$VEHICLES <- zap_labels(clean_pf$VEHICLES)
clean_pf$VEHICLES <- as.numeric(clean_pf$VEHICLES)
clean_pf$BEDROOMS <- zap_labels(clean_pf$BEDROOMS)
clean_pf$BEDROOMS <- as.numeric(clean_pf$BEDROOMS)
```

```{r}
clean_pf_categories <- clean_pf %>%
  mutate(Education = case_when(EDUC == 7 ~ "1st Year of College",
                               EDUC == 8 ~ "2nd Year of College",
                               EDUC == 9 ~ "3rd Year of College",
                               EDUC == 10 ~ "4th Year of College",
                               EDUC == 11 ~ "5th Year of College")) %>%
  mutate(Marital_Status = case_when(MARST == 1 | MARST == 2 ~ "Married",
                               MARST == 3 | MARST == 4 ~ "Divorced",
                               MARST == 5 ~ "Single",
                               MARST == "NA" ~ "Other"))
```



one hot encoding FOR marital status

```{r}
encoded_data_marital_status <- clean_pf_categories  %>%
  mutate(Married  = case_when(MARST == 1 | MARST == 2 ~ "1",
                               MARST != 1 | MARST != 2 ~ "0")) %>%
  mutate(Single  = case_when(MARST == 5 ~ "1",
                             MARST != 5 ~ "0")) %>%
  mutate(Divorced  = case_when(MARST == 3 | MARST == 4 ~ "1",
                             MARST != 3 | MARST != 4 ~ "0"))
```

```{r}
encoded_data_marital_status2 <- encoded_data_marital_status

encoded_data_marital_status2 <- clean_pf_categories  %>%
  mutate(Marital_Status = case_when(MARST == 1 | MARST == 2 ~ "Married",
                             MARST == 3 | MARST == 4 ~ "Divorced",
                             MARST == 5 ~ "Single",
                             MARST == "NA" ~ "Other")) %>%
  mutate(OWNERSHP_Rented  = case_when(OWNERSHP == 2  ~ "1",
                               OWNERSHP != 2  ~ "0")) %>%
  mutate(OWNERSHP_Owned  = case_when(OWNERSHP == 1 ~ "1",
                             OWNERSHP != 1 ~ "0")) %>%
  mutate(Fixed_Ownership  = case_when(OWNERSHP == 1 ~ "0", #so 0 means it's owned 
                             OWNERSHP == 2 ~ "1")) %>% #so 1 means it's rented %>%
  mutate(Married  = case_when(MARST == 1 | MARST == 2 ~ "1",
                               MARST != 1 | MARST != 2 ~ "0")) %>%
  mutate(Single  = case_when(MARST == 5 ~ "1",
                             MARST != 5 ~ "0")) %>%
  mutate(Divorced  = case_when(MARST == 3 | MARST == 4 ~ "1",
                             MARST != 3 | MARST != 4 ~ "0")) #%>%

```

```{r}
final_data <- encoded_data_marital_status2
```

```{r}
#Filter final data 

final_data <- final_data %>% 
  select(-c(Marital_Status,EDUCD,EDUC,MARST,OWNERSHP, Education))

# final_data <- final_data %>% 
#   select(-c(Education))

```

```{r}
glimpse(final_data)

final_data$VEHICLES <- as.factor(final_data$VEHICLES)
final_data$Fixed_Ownership <- as.factor(final_data$Fixed_Ownership)
final_data$Married <- as.factor(final_data$Married)
final_data$Single <- as.factor(final_data$Single)
# final_data$HighSchool <- as.factor(final_data$HighSchool)
# final_data$College <- as.factor(final_data$College)
#final_data$Single <- as.factor(final_data$Single)
# final_data$Add_College <- as.factor(final_data$Add_College)
final_data$BEDROOMS <- as.factor(final_data$BEDROOMS)
final_data$INCTOT <- as.factor(final_data$INCTOT)
final_data$AGE <- as.factor(final_data$AGE)

```

```{r}
final_data %>% drop_na()
#dim(final_data)

final_svm_draft <- final_data %>%
  select(c(INCTOT,AGE,BEDROOMS, VEHICLES,Fixed_Ownership,Married,Single,Divorced))

final_svm_draft %>% drop_na()
  
#Train test split
train <- sample(1:nrow(final_svm_draft), 800)
train.set <- final_svm_draft[train,]
test.set <- final_svm_draft[-train,]
```

```{r}
# p <- ggplot(final_data, aes(x=INCTOT, y=Fixed_Ownership))
# p + geom_point(size=3) +theme(text=element_text(size=18))
# 
# p <- ggplot(final_data, aes(x=INCTOT, y=Fixed_Ownership))
# p + geom_point(size=3) +theme(text=element_text(size=18))

```


Kernel = Linear 
```{r}
final.svm.linear <- svm(Fixed_Ownership ~., data = train.set, kernel = "linear", 
    cost = 0.01, scale = FALSE)
summary(final.svm.linear)
```

```{r}
pred.train <- predict(final.svm.linear, train.set)
cat('\nTraining Error rate is:')

1 - mean(pred.train == train.set$Fixed_Ownership)
table(predict = pred.train, truth = train.set$Fixed_Ownership)
```

Predict
```{r}
pred.test <- predict(final.svm.linear, test.set)
cat('\nTesting Error rate is:')

#Test err
1 - mean(pred.test == test.set$Fixed_Ownership)
table(predict = pred.test, truth = test.set$Fixed_Ownership)
```
I noticed that after randomly including Bedrooms out of pure interest, my results started to improve. 

Tuning 
```{r}
set.seed(1)
tune.out <- tune(svm, Fixed_Ownership ~ ., data = train.set, kernel = "linear", 
    ranges = list(cost = c(0.01, 0.1, 1, 3, 5, 8, 10)))

bestmodel <- tune.out$best.model
summary(bestmodel)
```

Compute training and test error for new value of cost, Cost = 1
```{r}
final.svm.linear.tuned <- svm(Fixed_Ownership ~ ., data = train.set, kernel = "linear", 
    cost = 1, scale = FALSE)

pred.train <- predict(final.svm.linear.tuned, train.set)
cat('\nTraining Error rate is:')

1 - mean(pred.train == train.set$Fixed_Ownership)
table(predict = pred.train, truth = train.set$Fixed_Ownership)

```

```{r}
pred.test <- predict(final.svm.linear.tuned, test.set)
cat('\nTesting Error rate is:')
1 - mean(pred.test == test.set$Fixed_Ownership)
table(predict = pred.test, truth = test.set$Fixed_Ownership)

final.svm.linear.tuned.confm <- table(predict = pred.test, truth = test.set$Fixed_Ownership)

```


```{r}
library(yardstick)
library(ggplot2)


# The confusion matrix from a single assessment set (i.e. fold)
cm.linear <- conf_mat(final.svm.linear.tuned.confm, obs, pred)

autoplot(cm.linear, type = "heatmap") +
  scale_fill_gradient(low="#F7AD50",high = "#3F97D0") +
  ggtitle("Confusion Matrix: Tuned Linear Kernel")
```


Kernel set to Radial

```{r}
final.svm.radial <- svm(Fixed_Ownership ~ ., data = train.set, kernel = "radial", 
    cost = 0.01, scale = FALSE)
summary(final.svm.radial)
```

Radial train/testing
```{r}
pred.train <- predict(final.svm.radial, train.set)
cat('\nTrain Error Rate:')
1 - mean(pred.train == train.set$Fixed_Ownership)
table(predict = pred.train, truth = train.set$Fixed_Ownership)
```

```{r}
pred.test <- predict(final.svm.radial, test.set)
cat('\nTesting Error rate is:')
1 - mean(pred.test == test.set$Fixed_Ownership)
table(predict = pred.test, truth = test.set$Fixed_Ownership)
```

tuning radial 
```{r}
#set.seed(1)
tune.out <- tune(svm, Fixed_Ownership ~ ., data = train.set, kernel = "radial", 
    ranges = list(cost = c(0.01, 0.1, 10, 25, 50),
                  gamma = c(0.5, 1, 2, 3, 4)))
summary(tune.out)

bestmodel <- tune.out$best.model
summary(bestmodel)
```

PLUG OUR VALUES in, cost = 10, gamma = 0.5

```{r}
final.svm.radial.tuned <- svm(Fixed_Ownership ~ ., data = train.set, kernel = 'radial',
             cost = 10, gamma = 0.5, scale = FALSE)

#Predict Train
pred.train <- predict(final.svm.radial.tuned, train.set)
cat('\nTraining Error rate is:')
1 - mean(pred.train == train.set$Fixed_Ownership)
table(predict = pred.train, truth = train.set$Fixed_Ownership)

#Predict Test
pred.test <- predict(final.svm.radial.tuned, test.set)
cat('\nTesting Error rate is:')
1 - mean(pred.test == test.set$Fixed_Ownership)
table(predict = pred.test, truth = test.set$Fixed_Ownership)


final.svm.radial.tuned.confm <- table(predict = pred.test, truth = test.set$Fixed_Ownership)
```

```{r}
library(yardstick)
library(ggplot2)


# The confusion matrix from a single assessment set (i.e. fold)
cm.radial <- conf_mat(final.svm.radial.tuned.confm, obs, pred)

autoplot(cm.radial, type = "heatmap") +
  scale_fill_gradient(low="#F7AD50",high = "#3F97D0") +
  ggtitle("Confusion Matrix: Tuned Radial Kernel")
```




Polynomial Kernel

```{r}
final.svm.poly <- svm(Fixed_Ownership ~ ., data = train.set, kernel = "polynomial", 
    cost = 1, degree = 2, scale = FALSE)
summary(final.svm.poly)
```

```{r}
pred.train <- predict(final.svm.poly, train.set)
cat('\nTrain Error Rate:')
1 - mean(pred.train == train.set$Fixed_Ownership)
table(predict = pred.train, truth = train.set$Fixed_Ownership)
```

Predict Test Set - Polynomial kernel
```{r}
pred.test <- predict(final.svm.poly, test.set)
cat('\nTesting Error rate is:')
1 - mean(pred.test == test.set$Fixed_Ownership)
table(predict = pred.test, truth = test.set$Fixed_Ownership)
```

tuning my polynomial kernel
```{r}
set.seed(1)
tune.out.poly <- tune(svm, Fixed_Ownership ~ ., data = train.set, kernel = "polynomial", 
    ranges = list(cost = c(0.01, 0.1, 1, 10, 25),
                  degree = c(1, 2, 3, 4, 5)))
summary(tune.out.poly)

bestmodel <- tune.out.poly$best.model
summary(bestmodel)
```

Now fit and predict using returned parameter cost = 0.01, degree = 1
```{r}
final.svm.poly.tuned <- svm(Fixed_Ownership ~ ., data = train.set, kernel = 'polynomial',
             cost = 0.01, degree = 1, scale = FALSE)

#Predict Train
pred.train <- predict(final.svm.poly.tuned, train.set)
cat('\nTraining Error rate is:')
1 - mean(pred.train == train.set$Fixed_Ownership)
table(predict = pred.train, truth = train.set$Fixed_Ownership)

#Predict Test
pred.test <- predict(final.svm.poly.tuned, test.set)
cat('\nTesting Error rate is:')
1 - mean(pred.test == test.set$Fixed_Ownership)
table(predict = pred.test, truth = test.set$Fixed_Ownership)


final.svm.poly.tuned.confm <- table(predict = pred.test, truth = test.set$Fixed_Ownership)
```

Gets the confusion matrix for Poly Kernel
```{r}
library(yardstick)
library(ggplot2)


# The confusion matrix from a single assessment set (i.e. fold)
cm.poly <- conf_mat(final.svm.poly.tuned.confm, obs, pred)

autoplot(cm.poly, type = "heatmap") +
  scale_fill_gradient(low="#F7AD50",high = "#3F97D0") +
  ggtitle("Confusion Matrix After Tuning Poly Kernel ")
```


```{r}
svm_final <- svm(Fixed_Ownership ~., data = train.set, kernel = "linear", 
    cost = 0.01, scale = FALSE)
summary(svm_final)
```


```{r}
# #Let's take it a sample size of 5000
# df_sample <- df[sample(nrow(final_data), 2000), ]
# 
# train <- sample(1:nrow(df_sample), nrow(df_sample) * 0.7)
# train.setg <- df_sample[train,]
# test.setg <- df_sample[-train,]
```

Compute the magnitude to get most significant variables

```{r}
# Train a linear SVM model with the final_data dataset

train.setg <- train.set
train.setg <- apply(train.setg, 2, as.numeric)


final.svm.linear.calc <- svm(Fixed_Ownership ~ ., data = train.setg, kernel = "linear")

# Get the coefficients for the linear SVM model
svm.coef <- coef(final.svm.linear.calc)

# Compute the magnitude of the coefficients
svm.coef.abs <- abs(svm.coef)

# Extract the top 3 input variables with the largest coefficient magnitudes
top.vars <- names(sort(svm.coef.abs, decreasing = TRUE)[1:8])
top.vars
```


final_svm_draft <- final_data %>%
  select(c(INCTOT,AGE, VEHICLES,Fixed_Ownership,Married,Single,Divorced))

final_svm_draft %>% drop_na()
  
#Train test split
train <- sample(1:nrow(final_svm_draft), 800)
train.set <- final_svm_draft[train,]
test.set <- final_svm_draft[-train,]


Produces svm plot for INCTOT and Age
```{r}
# split data into train and test sets
train.setg1 <- as.data.frame(train.setg)

training <- sample(1:nrow(train.setg1), size = 400)
train.plot <- train.setg1[training, ]
test.plot <- train.setg1[-training, ]

# create radial kernel SVM model
svm_rad_plot <- svm(Fixed_Ownership ~ INCTOT + AGE, data = train.plot, kernel = "radial")

# plot SVM decision boundary
plot(svm_rad_plot, train.plot, INCTOT~AGE,  
     main = 'SVM Radial Kernel Plot', 
     xlab = 'INCTOT', ylab = 'AGE',
     width = 10, height = 10)

# create grid of values for prediction
ind_income_grid <- seq(min(train.setg1$INCTOT), max(train.setg1$INCTOT), length.out = 100)
age_income_grid <- seq(min(train.setg1$AGE), max(train.setg1$AGE), length.out = 100)
grid_data <- expand.grid(INCTOT = ind_income_grid, AGE = age_income_grid)
predicted_data <- predict(svm_rad_plot, newdata = grid_data)

# combine prediction with grid data
plot_data <- cbind(grid_data, prediction = predicted_data)

# create SVM plot
# create SVM plot
ggplot(data = plot_data, aes(x = INCTOT, y = AGE)) +
  geom_point(aes(color = prediction)) +
  scale_color_gradient(low = "red", high = "blue") +
  stat_contour_filled(aes(z = as.numeric(prediction)), alpha = 0.2) +
  labs(title = "SVM Radial Kernel Plot", x = "Individual Income", y = "Age") +
  theme_bw() + guides(fill = FALSE) #Remove Legends 
```

Produces svm plot for INCTOT and Bedrooms
```{r}
# split data into train and test sets
train.setg1 <- as.data.frame(train.setg)

training <- sample(1:nrow(train.setg1), size = 400)
train.plot <- train.setg1[training, ]
test.plot <- train.setg1[-training, ]

# create radial kernel SVM model
svm_rad_plot <- svm(Fixed_Ownership ~ INCTOT + BEDROOMS, data = train.plot, kernel = "radial")

# plot SVM decision boundary
plot(svm_rad_plot, train.plot, INCTOT~AGE,  
     main = 'SVM Radial Kernel Plot', 
     xlab = 'INCTOT', ylab = 'BEDROOMS',
     width = 10, height = 10)

# create grid of values for prediction
ind_income_grid <- seq(min(train.setg1$INCTOT), max(train.setg1$INCTOT), length.out = 100)
bedrooms_grid <- seq(min(train.setg1$BEDROOMS), max(train.setg1$BEDROOMS), length.out = 100)
grid_data <- expand.grid(INCTOT = ind_income_grid, BEDROOMS = bedrooms_grid)
predicted_data <- predict(svm_rad_plot, newdata = grid_data)

# combine prediction with grid data
plot_data <- cbind(grid_data, prediction = predicted_data)

# create SVM plot
# create SVM plot
ggplot(data = plot_data, aes(x = INCTOT, y = BEDROOMS)) +
  geom_point(aes(color = prediction)) +
  scale_color_gradient(low = "red", high = "blue") +
  stat_contour_filled(aes(z = as.numeric(prediction)), alpha = 0.2) +
  labs(title = "SVM Radial Kernel Plot", x = "Individual Income", y = "Bedrooms") +
  theme_bw() + guides(fill = FALSE) #Remove Legends 
```


Produces svm plot for Bedrooms and Age
```{r}

# split data into train and test sets
train.setg1 <- as.data.frame(train.setg)

training <- sample(1:nrow(train.setg1), size = 400)
train.plot <- train.setg1[training, ]
test.plot <- train.setg1[-training, ]

# create radial kernel SVM model
svm_rad_plot <- svm(Fixed_Ownership ~ AGE + BEDROOMS, data = train.plot, kernel = "radial")

# plot SVM decision boundary
plot(svm_rad_plot, train.plot, AGE~BEDROOMS,  
     main = 'SVM Radial Kernel Plot', 
     xlab = 'AGE', ylab = 'BEDROOMS',
     width = 10, height = 10)

# create grid of values for prediction
age_grid <- seq(min(train.setg1$AGE), max(train.setg1$AGE), length.out = 100)
bedrooms_grid <- seq(min(train.setg1$BEDROOMS), max(train.setg1$BEDROOMS), length.out = 100)
grid_data <- expand.grid(AGE = age_grid, BEDROOMS = bedrooms_grid)
predicted_data <- predict(svm_rad_plot, newdata = grid_data)

# combine prediction with grid data
plot_data <- cbind(grid_data, prediction = predicted_data)

# create SVM plot
# create SVM plot
ggplot(data = plot_data, aes(x = AGE, y = BEDROOMS)) +
  geom_point(aes(color = prediction)) +
  scale_color_gradient(low = "red", high = "blue") +
  stat_contour_filled(aes(z = as.numeric(prediction)), alpha = 0.2) +
  labs(title = "SVM Radial Kernel Plot", x = "Age", y = "Bedrooms") +
  theme_bw() + guides(fill = FALSE) #Remove Legends 
```


```{r}
ggplot(final_data, aes(x = Fixed_Ownership)) +
  geom_bar(fill = "#F5DEB3") +
  labs(x = "Ownership", y = "Count") +
  ggtitle("Ownership Distribution") +
  scale_fill_manual(values = c("owned" = "#F5DEB3", "rent" = "#F5DEB3")) +
  theme_void() +
  theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14, angle = 90),
        axis.text = element_text(size = 14),
        axis.line = element_line(color = "#2E2E2E")) +
  scale_x_discrete(labels = c("0" = "Owns", "1" = "Rents"))
```








